---
title: "Notes"
author: "Salim M'jahad"
output: pdf_document
---

# Types
## Numeric
```{r}
x <- 2
mode(x)
```
```{r}
typeof(x)
```
```{r}
y <- as.integer(2)
typeof(x)
```

## Complex
```{r}
z <- 2 + 5i
mode(z)
```
```{r}
typeof(z)
```

## Character
```{r}
name <- "Something :)"
typeof(name)
```

## Boolean
```{r}
boolT <- TRUE
boolF <- F
typeof(boolT)
```

## Data Types:
\newline $\bullet$ Vectors
\newline $\bullet$ Scalars
\newline $\bullet$ Matrices
\newline $\bullet$ Arrays
\newline $\bullet$ Lists (can have different modes)
\newline $\bullet$ Dataframes (can have different modes)



# Functions

## Building Vectors
```{r}
5:10
rep("hello", 5)
v <- c()
v[1] <- 6
v[3] <- 5
v
```

## Concatenate
```{r}
c(c(1,2), c(3,4,5))
```

## Matrix
```{r}
matrix(1:9,nrow=3, ncol=3)
matrix(1:9,nrow=3, ncol=3, byrow=T)
```

```{r}
lol <- matrix(nrow=2, ncol=3)
lol[1,2] <- 3
rbind(lol, c(4:6))
```

## Help

```{r}
?str
??"hilbert matrix" 
```

## Subsets

```{r}
tmp <- c(11, 12, 13, 14, 15)
tmp[2:4]
tmp[c(1,4)]
tmp[c(2,5)] <- 3
tmp
tmp[-1]
tmp[-c(1,4)]
```


```{r}
tmp <- matrix(1:6, nrow=2, ncol=3)
tmp[,2:3]
tmp[1,]
```

## Installing Packages
```{r}
#install.packages("ggplot2")
#install.packages("pixmap")
```

## Importing Packages
```{r}
library("ggplot2")
library("pixmap")
```

## Manipulating images as matrices
```{r}
bird <- read.pnm("bird.pgm")
plot(bird)
```

```{r}
bird@grey[100:200, 200:300] <- 1
bird@grey[150:250, 250:350] <- 0
plot(bird)
```

## Matrix Operations
```{r}
mat1 <- rbind(c(1,2),c(4,3))
mat2 <- 3*mat1
mat2 == mat1 + mat1 + mat1

transpose <- t(mat2)
determinant <- det(mat2)
inverse <- solve(mat2)
eigenvals <- eigen(mat2)$val
eigenvecs <- eigen(mat2)$vec
svd_D <- svd(mat2)$d
```

Other functions:
\newline $\bullet$ length(x) : Length of a vector x
\newline $\bullet$ sum(x) : Sum of a vector x
\newline $\bullet$ mean(x) : Arithmetic mean of a vector x
\newline $\bullet$ quantiles(x) : Sample quantiles of a vector x 
\newline $\bullet$ max(x) : Maximum of a vector x
\newline $\bullet$ min(x) : Minimum of a vector x
\newline $\bullet$ sd(x) : Standard deviation of a vector x
\newline $\bullet$ var(x) : Variance of a vector x
\newline $\bullet$ summary(x) : Summary statistics of vector x

## Factors

### Split()
```{r}
group <- factor(rep(c("Control","Treatment"),c(3,4)))
ages <- c(20, 30, 40, 35, 35, 35, 35)
sex <- c("M", "M", "F", "M", "F", "F", "F")
split(ages, list(group, sex))
```

### Table() for Contigency Tables
```{r}
table(sex, group)
```


### Plot x=Frost y=Illiteracy from dataset=States
```{r}
#plot(Illiteracy ~ Frost, data = states)
```


### Read .txt data
read.table("stu_data.txt", header=FALSE, as.is=TRUE)
(you can specificy the seperation like commas with sep= )
scan() reads anything into a vector with elements separated by whitespace


### Alphabetize a list of vectors with each elements represented by a word (names(wordlist))
```{r}
alphabetized_list <- function(wordlist) {
   nms <- names(wordlist) # The names are the words
   sorted <- sort(nms) # The words, but now in ABC order
   return(wordlist[sorted]) # Returns the sorted version +}
}
```

## Loops
```{r}
smtg <- c(1,2,3)
for (i in smtg) {
  print(i)
}
```

```{r}
tapply(ages, group, mean)
```

## Exploratory Data Analysis

\newline 1. Generate questions about your data.
\newline 2. Search for answers by visualizing, transforming, and modelling your
data.
\newline 2. Use what you learn to refine your questions and or generate new questions.



### Number of Character
nchar("hello") == 5
length() number of strings in the object

substr(string, start = , stop = )

strsplit(todo, split = ", ")

paste("cat", "squirrel", "hedgehog", sep = ", ")
paste(c("cat", "squirrel", "hedgehog"), 1:3)







