---
title: "Lab 5"
author: "Salim M'jahad msm2243"
date: "February 15, 2018"
output: pdf_document
---

# Part 1

## i. 
```{r}
housing = read.csv("NYChousing.csv")
```

## ii.
```{r}
dim(housing)
```
The dataframe has 2506 rows and 22 columns

## iii. 
```{r}
apply(is.na(housing), 2, sum)
```

This function applies the function sum on the values in housing that are NA and returns the results in a vector format.

## iv. 
```{r}
housing <- subset(housing, !is.na(Value))
```


## v.
```{r}
dim(housing)
```

2454 = 2506 - 52

I removed 52 rows, so yes it agrees.

## vi.
```{r}
housing$logValue <- log(housing$Value)
summary(housing$logValue)
```
minimum = 8.41
median = 13.75
mean = 13.68
maximum = 20.47


## vii.
```{r}
housing$logUnits <- log(housing$UnitCount)
summary(housing$logUnits)
```

## viii.
```{r}
housing$after1950 <- housing$YearBuilt >= 1950
summary(housing$after1950)
```

# Part 2

## i.
```{r}
library(ggplot2)

p1 <- ggplot(data = housing) + geom_point(mapping = aes(y = logValue, x = logUnits))
p1
```

## ii.
```{r}
p2 <- ggplot(data = housing) + geom_point(mapping = aes(y = logValue, x = logUnits, col=after1950)) 
p2
```

There is almost a direct linear relationship between the logUnits variable and the logValue variable. The more units there are in a property the larger its value. The covariance seems to be positive.

## iii.
```{r}
man <- subset(housing, Borough=="Manhattan")
brook <- subset(housing, Borough=="Brooklyn")
a1950 <- subset(housing, after1950 == TRUE)
b1950 <- subset(housing, after1950 == FALSE)

cov_all <- cov(x=housing$logValue, y=housing$logUnits)
cov_man <- cov(x=man$logValue, y=man$logUnits)
cov_brook <- cov(x=brook$logValue, y=brook$logUnits)
cov_a1950 <- cov(x=a1950$logValue, y=a1950$logUnits)
cov_b1950 <- cov(x=b1950$logValue, y=b1950$logUnits)

c(cov_all, cov_man, cov_brook, cov_a1950, cov_b1950)
```
(i) the whole data: 2.182148
(ii) just Manhattan: 1.983556
(iii) just Brooklyn: 2.566733
(iv) for properties built after 1950: 1.123973
(v) for properties built before 1950: 1.519853

## iv.

```{r}
manorbrook <- subset(housing, Borough %in% c("Manhattan", "Brooklyn"))
p3 <- ggplot(data = manorbrook) + geom_point(mapping = aes(y = logValue, x = logUnits, col =Borough)) 
p3
```

## v.
```{r}
median(subset(housing, Borough=="Manhattan")$Value, na.rm = TRUE)
```

## vi.

```{r}
p4 <- ggplot(data=housing) + geom_boxplot(aes(x=factor(Borough), y=logValue))
p4
```

## vii.

```{r}
X <- split(housing, housing$Borough)

medi <- function(datf) {
  return(median(datf$Value, na.rm = TRUE))
}
sapply(X, medi)
```